---
title: "PLSR_clean"
author: "Hannah Curtis"
date: "2024-05-06"
output: html_document
---

Running PLSR for green infrastructure data using find_best_model algorithm that selects variables based on VIP scores/AIC values. Response variable is infiltration rate of 182 biofilters in WI and surrounding states. 

Updated: 7/15/24

read in variable file
```{r}
library(vip)
library(caret)
library("corrr")
library(ggcorrplot)
library("FactoMineR")
library("factoextra")
library("car")
library("pls")
library("plsdepot")
library(ggplot2)
library(tidyr)
library(dplyr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(stats)
library(corrplot)

df_all <- read.csv("lm_data_all.csv")
df_pred_removed <- read.csv("lm_data_na_predictors_removed.csv")
df_res_removed <- read.csv("lm_data_na_responses_removed.csv")
df_subset <- read.csv("lm_data_all_subset.csv")
df_subset_res_removed <- read.csv("lm_data_all_subset_res_removed.csv")
```

Correlation Matrix
```{r}
cor_matrix <- cor(df_res_removed[4:44])
corrplot(cor_matrix, method = "color",  tl.col = "black", tl.srt = 45, tl.cex = 0.5,title = "Correlation Matrix")
```

PCA of predictor variables--Add PCs to data frame
```{r}
# Run PCA
# Areas
pca_area <- prcomp(df_res_removed[4:6], scale=TRUE, center=TRUE)
summary(pca_area)
# Extract PCA1
first_pc_area <- pca_area$x[, 1]

# Watershed Land Use
pca_landuse <- prcomp(df_res_removed[12:15], scale=TRUE, center=TRUE)
summary(pca_landuse)
# Extract PCA1
first_pc_landuse <- pca_landuse$x[, 1]

# Vegetation Type
pca_veg <- prcomp(df_res_removed[16:21], scale=TRUE, center=TRUE)
summary(pca_veg)
# Extract PCA1
first_pc_veg <- pca_veg$x[, 1]

# Underlying Soil
pca_soil <- prcomp(df_res_removed[22:27], scale=TRUE, center=TRUE)
summary(pca_soil)
# Extract PCA1
first_pc_soil <- pca_soil$x[, 1]

# Engineered Soil
pca_engsoil <- prcomp(df_res_removed[39:44], scale=TRUE, center=TRUE)
summary(pca_engsoil)
# Extract PCA1
first_pc_engsoil <- pca_engsoil$x[, 1]

# Poor Infiltration
pca_poorinf <- prcomp(df_res_removed[28:29], scale=TRUE, center=TRUE)
summary(pca_poorinf)
# Extract PCA1
first_pc_poorinf <- pca_poorinf$x[, 1]

# Inlet Type
pca_inlet <- prcomp(df_res_removed[31:34], scale=TRUE, center=TRUE)
summary(pca_inlet)
# Extract PCA1
first_pc_inlet <- pca_inlet$x[, 1]

# Outlet Type
pca_outlet <- prcomp(df_res_removed[35:38], scale=TRUE, center=TRUE)
summary(pca_outlet)
# Extract PCA1
first_pc_outlet <- pca_outlet$x[, 1]

# Bind PCA1 to dataframe
df_with_pc <- cbind(df_res_removed, Area_PC = first_pc_area, LandUse_PC = first_pc_landuse, Veg_PC = first_pc_veg, Soil_PC = first_pc_soil, EngSoil_PC = first_pc_engsoil, PoorInf_PC = first_pc_poorinf, Inlet_PC = first_pc_inlet, Outlet_PC = first_pc_outlet )
```

Calculate z-scores and remove outliers (may want to run this for predictor variables to check outliers)
```{r}
# Write a function to calculate z-scores for each variable in the df_storm dataframe and remove any row of the data frame where the z-score is above 3 or below -3
# I only want to calculate the z-scores of certain variables in the dataframe when considering outliers
remove_outliers <- function(df) {
  # Calculate z-scores for each variable in the dataframe
  z_scores <- scale(df[, c("Area",	"Contributing.area",	"Drainage.Area.Ratio",	"Age",	"Ponding.depth",	"Inlet.condition",	"Outlet.condition",	"Underdrain", "X..Residential", "X..Commercial", "X..Industrial", "X..Institutional", "X..Vegetation_shrub",	"X..Vegetation_tree",	"X..Vegetation_grass",	"X..Vegetation_prairie",	"X..Vegetation_bare",  "X..Vegetation_other",	"Underlying.soil_..sand",	"Underlying.soil_..silt",	"Underlying.soil_..clay",	"Underlying.soil_..urban.disturbed",	"Underlying.soil_..other.undefined",	"Underlying.soil_..organic.matter",	"Evidence.of.poor.infiltration",	"X..of.surface.with.poor.infiltration",	"Vegetation.condition",	"Inlet.feature_Curb.Cut",	"Inlet.feature_Overland.Flow",	"Inlet.feature_Pipe",	"Outlet.feature_Overland.Flow", "Outlet.feature_Pipe",	"Outlet.feature_Riser.Pipe",	"Engineered.soil.depth",	"Engineered.soil._..sand",	"Engineered.soil._..silt",	"Engineered.soil._..clay",	"Engineered.soil._..compost",	"Engineered.soil._..other")])
  
  # Remove any row of the data frame where the z-score is above 3 or below -3
  df <- df[apply(z_scores, 1, function(x) all(abs(x) < 3)), ]
  
  return(df)
}

df_pred_outliers_removed <- remove_outliers(df_res_removed)
```

Remove Infiltration Rate Outliers 
```{r}
# Write a function to calculate z-scores for each variable in the df_storm dataframe and remove any row of the data frame where the z-score is above 3 or below -3
# I only want to calculate the z-scores of certain variables in the dataframe when considering outliers
inf_outliers <- function(df) {
  # Calculate z-scores for each variable in the dataframe
  z_scores <- scale(df[, c("Average.infiltration.rate")])
  
  # Remove any row of the data frame where the z-score is above 3 or below -3
  df <- df[apply(z_scores, 1, function(x) all(abs(x) < 3)), ]
  
  return(df)
}

df_outliers_removed <- inf_outliers(df_res_removed)
```

Infiltration Rate:
```{r}
source("find_best_model_GI.R")
df <- df_with_pc

y = c("Average.infiltration.rate")
X = c("Area",	"Contributing.area",	"Drainage.Area.Ratio",	"Age",	"Ponding.depth",	"Inlet.condition",	"Outlet.condition",	"Underdrain", "X..Residential", "X..Commercial", "X..Industrial", "X..Institutional", "X..Vegetation_shrub",	"X..Vegetation_tree",	"X..Vegetation_grass",	"X..Vegetation_prairie",	"X..Vegetation_bare",  "X..Vegetation_other",	"Underlying.soil_..sand",	"Underlying.soil_..silt",	"Underlying.soil_..clay",	"Underlying.soil_..urban.disturbed",	"Underlying.soil_..other.undefined",	"Underlying.soil_..organic.matter",	"Evidence.of.poor.infiltration",	"X..of.surface.with.poor.infiltration",	"Vegetation.condition",	"Inlet.feature_Curb.Cut",	"Inlet.feature_Overland.Flow",	"Inlet.feature_Pipe",	"Outlet.feature_Overland.Flow", "Outlet.feature_Pipe",	"Outlet.feature_Riser.Pipe",	"Engineered.soil.depth",	"Engineered.soil._..sand",	"Engineered.soil._..silt",	"Engineered.soil._..clay",	"Engineered.soil._..compost",	"Engineered.soil._..other","Area_PC", "LandUse_PC", "Veg_PC", "Soil_PC", "PoorInf_PC", "Inlet_PC", "Outlet_PC", "EngSoil_PC")
output_df <- get_best_vars_generic(X,y,df,reduce_model="vi", overall_obj="aic")
output_df

```
Testing PLSR
```{r}

# make this example reproducible
set.seed(0)

# Predictor vars candidates to add back in: 
# Area+Contributing.area+Drainage.Area.Ratio+Age+Ponding.depth+Inlet.condition+Outlet.condition+Underdrain+X..Residential+X..Commercial+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..clay+Underlying.soil_..urban.disturbed+Underlying.soil_..other.undefined+Underlying.soil_..organic.matter+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Vegetation.condition+Inlet.feature_Curb.Cut+Inlet.feature_Other+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Other+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other

# Categorical vars: Underdrain+Evidence.of.poor.infiltration+Inlet.feature_Curb.Cut+Inlet.feature_Other+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Other+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe

model <- plsr(Average.infiltration.rate ~ 	
Area+Contributing.area+Drainage.Area.Ratio+Age+Ponding.depth+Inlet.condition+Outlet.condition+Underdrain+X..Residential+X..Commercial+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..clay+Underlying.soil_..urban.disturbed+Underlying.soil_..other.undefined+Underlying.soil_..organic.matter+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Vegetation.condition+Inlet.feature_Curb.Cut+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other+Area_PC +LandUse_PC+Veg_PC+Soil_PC +PoorInf_PC+Inlet_PC+Outlet_PC+EngSoil_PC 
,
  data=df_with_pc, scale=TRUE, validation="CV", segments=5,  ncomp=4
)

# RMSEP and % variance explained
summary(model)

# Extract coefficients
coefficients <- coef(model)

# visualize cross-validation plots
validationplot(model)
validationplot(model, val.type="MSEP")
#validationplot(model, val.type="R2")

# Make plots square (better to visualize predictions when x and y axes are the same)
par(pty="s")

# Plot predicted vs measured using first 3 components
plot(model, ncomp=4, line=TRUE, main="",  xlab="Measured Infiltration Rate", ylab="Predicted Infiltration Rate", pch = 16, col = "salmon", bg="transparent", cex.axis=1)
grid()

# R-squared
print(pls::R2(model))
#print(Q2(model))

vi(model, ncomp=8)
coefs <- coef(model)

# Predict the response variable
predicted_values <- predict(model)

# AIC
#get_aic(model)

```

Linear Regressions:
```{r}
x <- df_res_removed$Drainage.Area.Ratio
  
poly_model <- lm(df_res_removed$Average.infiltration.rate ~ poly(x,2))
plot(x, df_res_removed$Average.infiltration.rate, pch = 16, col = "seagreen3", xlab="Drainage Area Ratio", ylab="Average Infiltration Rate", bg="transparent", cex.axis=1)
grid()

summary(poly_model)

#use model to get predicted values
pred <- predict(poly_model)
ix <- sort(x, index.return=T)$ix

#add polynomial curve to plot
lines(x[ix], pred[ix], col='black', lwd=2)
```


Bar Charts
```{r}
# Assuming infil_vip is your vector of VIP scores
infil_vip_df <- data.frame(variable = infil_vip$Variable, VIP_score = infil_vip$Importance)

# Arrange data frame from highest to lowest VIP scores
#infil_vip_df %>% arrange(desc(VIP_score))
infil_vip_df_desc <- arrange(infil_vip_df, -VIP_score)

# Add variable type as column to data frame
# Create a vector of variable types corresponding to each variable
#variable_types <- c("HYD", "ED", "WS", "HYD", "HYD", "WS", "ED", "ED", "ED", "HYD", "ED", "HYD", "HYD", "HYD", "ED", "ED")  # Adjust with your actual variable types

# Add alphabetical column to data frame so that bar chart will be plotted correctly
# Create an alphabetical vector
alphabet <- c("V", "U", "T", "S", "R", "Q", "P", "O", "N", "M", "L", "K", "J", "I", "H", "G", "F", "E", "D", "C", "B", "A")

# Add the variable type column to infil_vip_df
#infil_vip_df$type <- variable_types

# Sort the data frame by variable type
#infil_vip_df %>% arrange(type)
#infil_vip_df_grouped <- arrange(infil_vip_df, type)

# Add alphabetical column to data frame
infil_vip_df$alphabet <- alphabet

# Create custom labels for y axis that are the same as the infil_vip_df_grouped$variable column and in the same order
custom_labels <- c("% Vegetation Other", "Engineered Soil % Compost", "% Vegetation Prairie", "Drainage Area Ratio", "Engineered Soil % Clay", "Pipe Inlet Feature", "% of Surface with Poor Infiltration", "% Vegetation Shrub", "Evidence of Poor Infiltration", "% Industrial", "Underlying Soil % Silt", "Underlying Soil % Sand", "Underdrain", "% Vegetation Grass", "% Residential", "Engineered Soil % Other", "Inlet Condition", "Pipe Outlet Feature", "Overland Flow Outlet Feature", "Riser Pipe Outlet Feature", "Overland Flow Inlet Feature", "% Vegetation Tree")
 

# Create a custom_labels vector that is the same as above but switch the order of all of the labels so the last one above is the first in this vector, etc.
custom_labels <- rev(custom_labels)

# Plot the sideways bar chart by alphabet column to get the variables grouped by type and sorted by VIP score within each type
ggplot(infil_vip_df_desc, aes(x = VIP_score, y = alphabet, fill = VIP_score)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "white", high = "limegreen", name = "VIP Score") +
  scale_y_discrete(labels = custom_labels) +
  labs(x = "VIP Score", y = "") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, size = 12)) 
  #annotate(geom="text", x=1.6, y=1.5, label="Hydrologic Variables")
```

All vars: Area+Contributing.area+Drainage.Area.Ratio+Age+Ponding.depth+Inlet.condition+Outlet.condition+Underdrain+X..Residential+X..Commercial+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..clay+Underlying.soil_..urban.disturbed+Underlying.soil_..other.undefined+Underlying.soil_..organic.matter+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Vegetation.condition+Inlet.feature_Curb.Cut+Inlet.feature_Other+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Other+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other

"Area",	"Contributing.area",	"Drainage.Area.Ratio",	"Age",	"Ponding.depth",	"Inlet.condition",	"Outlet.condition",	"Underdrain", "X..Residential", "X..Commercial", "X..Industrial", "X..Institutional", "X..Vegetation_shrub",	"X..Vegetation_tree",	"X..Vegetation_grass",	"X..Vegetation_prairie",	"X..Vegetation_bare",  "X..Vegetation_other",	"Underlying.soil_..sand",	"Underlying.soil_..silt",	"Underlying.soil_..clay",	"Underlying.soil_..urban.disturbed",	"Underlying.soil_..other.undefined",	"Underlying.soil_..organic.matter",	"Evidence.of.poor.infiltration",	"X..of.surface.with.poor.infiltration",	"Vegetation.condition",	"Inlet.feature_Curb.Cut", "Inlet.feature_Other",	"Inlet.feature_Overland.Flow",	"Inlet.feature_Pipe", "Outlet.feature_Other",	"Outlet.feature_Overland.Flow", "Outlet.feature_Pipe",	"Outlet.feature_Riser.Pipe",	"Engineered.soil.depth",	"Engineered.soil._..sand",	"Engineered.soil._..silt",	"Engineered.soil._..clay",	"Engineered.soil._..compost",	"Engineered.soil._..other"

Have to delete: Inlet.feature_Other, Outlet.feature_Other

Hand Selected Vars:Drainage.Area.Ratio+Age+Ponding.depth+Underdrain+X..Residential+X..Commercial+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..clay+Underlying.soil_..urban.disturbed+Underlying.soil_..other.undefined+Underlying.soil_..organic.matter+X..of.surface.with.poor.infiltration+Vegetation.condition+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other


VI, Q2: Drainage.Area.Ratio+Inlet.condition+Underdrain+X..Residential+X..Industrial+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other

VI_AIC, Q2: Area+Drainage.Area.Ratio+Ponding.depth+Inlet.condition+Underdrain+X..Residential+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..other.undefined+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other

VI, AIC: Drainage.Area.Ratio+X..Vegetation_shrub+X..Vegetation_other+Engineered.soil._..compost

VI_AIC, AIC: Drainage.Area.Ratio+X..Vegetation_shrub+X..Vegetation_other+Engineered.soil._..compost

Best manual model: X..Residential+Drainage.Area.Ratio+X..Vegetation_shrub+X..Vegetation_other+Engineered.soil._..compost+X..Vegetation_tree+X..Vegetation_prairie+Underlying.soil_..sand+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Outlet.feature_Pipe+X..Vegetation_grass

Response outliers removed: Area+Drainage.Area.Ratio+Inlet.condition+Outlet.condition+Underdrain+X..Commercial+X..Vegetation_shrub+X..Vegetation_other+Underlying.soil_..sand+Evidence.of.poor.infiltration+Outlet.feature_Pipe+Engineered.soil.depth+Engineered.soil._..compost

Predictor vars removed: Contributing.area+Drainage.Area.Ratio+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..silt+Engineered.soil._..compost+Engineered.soil._..other